---
title: " A Data-Driven Analysis of Football Player Performance and Wage"
author: "Olga Litvinova"
editor: visual
format: 
  html:
    fig-height: 7
    fig-width: 7
    fig-format: png
    fig-dpi: 300
---

**Football** is one of the most popular sports in the world, enjoyed by billions of fans and generating huge amounts of money through TV rights, sponsorship, and merchandise. Because of this, football players often earn some of the highest salaries in professional sports. Clubs spend a lot of money on players based on their skills, physical abilities, and overall performance.

This project is divided into two major components:

1.  **EDA** - We will start with exploratory data analysis, and see the connection between player performance and salary to understand what factors influence how much footballers get paid.\
2.  **Regression model** - Then, we will use regression modeling to see how different performance metrics affect player wages, giving us insights into what really drives player value in football.For that we will apply Recursive Feature Elimination (RFE) and backward elimination from ML.

By the way, to ensure a consistent and meaningful analysis, we focus exclusively on players from the **top 10 football leagues**. The league rankings used in this analysis were sourced from [Football Flick's Top 10 Leagues](https://footballflick.com/en-gr/blogs/news/the-best-leagues-in-the-world-top-10-ranked).

For the purposes of this study, **Turkey's Süper Lig** was included in place of **USA/Canada**.

Additionally, goalkeepers are excluded from the analysis because their performance metrics differ significantly from those of outfield players, and including them could introduce bias into the results.

```{r, warning=FALSE, message=FALSE, echo = FALSE}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(ggrepel)
library(scales)
library(plotly)
library(hrbrthemes)
library(showtext)

###Download Fonts
#must download font awesome locally first! https://fontawesome.com/download
showtext.auto()
setwd("C:\\Users\\Olga\\0. VS Code\\Github\\visualizations\\")
sysfonts::font_add('fs', 'fonts/otfs/Font Awesome 6 Free-Solid-900.otf')
sysfonts::font_add('fb', 'fonts/otfs/Font Awesome 6 Brands-Regular-400.otf')
sysfonts::font_add_google("Open Sans","opensans")
font_sans<-"opensans"
#sysfonts::font_add_google("IBM Plex Sans", "ibm_plex_sans")
font_add_google("IBM Plex Sans", "ibm_plex_sans_bold", bold.wt = 700)

#Setting my theme
theme_ipsum_rc <-theme_ipsum_rc(base_family = "ibm_plex_sans")+
  theme(
    text=element_text(size=14, color="#333333"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    plot.title = element_text(size=16, hjust = 0,vjust = 1,family = "fs", face = "bold", color="666666"),
    axis.text = element_text(size = 14,family = "ibm_plex_sans"),
    geom_text =element_text(size = 14,family = "ibm_plex_sans"),
    panel.grid.minor=element_blank(),
    panel.grid.major=element_line(color="gray90"),
    legend.position="top"
  )
theme_set(theme_ipsum_rc)


```

## Data

The dataset used in this project was sourced from [SoFIFA](https://sofifa.com/players?type=all&col=wg&sort=desc), a platform that provides detailed information on football players from various leagues worldwide. The data reflects player statistics as of **the end of 2024**

The dataset includes the following key information:

-   **Personal Details:** Player name, age, preferred foot, height, and weight.\
-   **Club & Contract:** League, contract duration, release clauses.\
-   **Performance Metrics:** Technical skills (e.g., finishing, dribbling), physical attributes (e.g., sprint speed, stamina), and mental attributes (e.g., vision, composure).\
-   **Market Value:** Estimated player value and salary information.\
-   **Positional Data:** Best position and player role (attacker, midfielder, defender, goalkeeper).

In this analysis, goalkeepers have been excluded to ensure consistency and accuracy when examining player performance, market value, and salary. This decision was made because goalkeepers have unique performance metrics (such as GK.Reflexes, GK.Diving, GK.Handling, and GK.Positioning) that are not comparable to the metrics used for outfield players, such as Finishing, Dribbling, Passing, and Sprint Speed.

```{r, warning=FALSE, message=FALSE, echo = FALSE}
setwd("C:\\Users\\Olga\\0. VS Code\\Github\\visualizations\\")

df<-read.csv("data\\fifa_players.csv")%>%
  filter(Source_File=="combined_fifa_players_20241112_163726")
```

### Data Manipulation

```{r, warning=FALSE, message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
library(dplyr)
league_rankings <- data.frame(
  League = c( "Süper Lig (Türkiye)", 
             "Eredivisie (Netherlands)","Primeira Liga (Portugal)","Liga Profesional de Fútbol (Argentina)", "Série A (Brazil)",  "Ligue 1 (France)",
             "Serie A (Italy)", "Bundesliga (Germany)", "La Liga (Spain)", "Premier League (England)"),
  League_Rank = 10:1  
)
numeric_vars <- c("Finishing", "Dribbling", "Shot.Power", "Sprint.Speed", 
                  "Attack.Position", "Short.Passing", "Vision", "Long.Passing", 
                  "Stamina", "Standing.Tackle", "Interceptions", "Strength", 
                  "Defensive.Awareness", "Heading.Accuracy", "Crossing","Overall.Score", 
                  "Potential.Score", "Composure", "Reactions","International.Reputation", "Aggression",
                  "Total.Movement", "Total.Power", "Total.Defending", "Total.Mentality", "Total.Stats" ,"Volleys","Curve","FK.Accuracy","Ball.Control", "Acceleration","Agility","Balance","Jumping" ,"Long.Shots", "Penalties","Sliding.tackle")


df_clean<- df %>% 
  left_join(league_rankings, by = "League") %>% 
  mutate(
    Wage = as.numeric(str_replace_all(str_remove_all(Wage, "€"), c("K" = "000", "M" = "000000"))),
    Value = as.numeric(str_replace_all(str_remove_all(Value, "€"), c("K" = "000", "M" = "000000"))),
    Contract.Start = as.numeric(Contract.Start),
    Contract.End = as.numeric(Contract.End),
    Contract.Duration=Contract.End-Contract.Start,
    Height_cm = as.numeric(str_extract(Height, "\\d+")),
    Weight_kg = as.numeric(str_extract(Weight, "\\d+")),
    Left_Foot = ifelse(Preferred.Foot == "Left", 1, 0),
    across(all_of(numeric_vars), 
                ~ as.numeric(str_remove(., "\\+\\d+"))),
    Role = case_when(
      #Option1
      str_detect(Best.Position, "LW|RW") ~ "Winger",
      str_detect(Best.Position, "ST|CF") ~ "Striker",
      str_detect(Best.Position, "CM|CAM|CDM") ~ "Central Midfielder",
      str_detect(Best.Position, "CB") ~ "Central Defender",
      str_detect(Best.Position, "RWB|LWB|LB|RB") ~ "Side Defender",
      str_detect(Best.Position, "LM|RM") ~ "Side Midfielder",
      TRUE ~ "Unknown"  
  ),
  LeagueRanked = paste0("#", League_Rank, " ", League)
  ) %>%
  filter(Role != "Unknown", !is.na(Overall.Score), !is.na(Wage),   !is.na(Contract.Duration), League %in% league_rankings$League, Best.Position != "GK")%>%
    dplyr::select(
    -c(Position, Player.ID, Loan.End, Attacking.Work.Rate, Defensive.Work.Rate, 
       Source_File, Joined.Team, Height, Weight, Real.Face, Body.Type, 
       Contract.Start, Contract.End, Total.Goalkeeping,Preferred.Foot,Best.Overall,Release.Clause, Weak.Foot,Best.Position),
    -matches("^(GK\\.|Total\\.)")
  )
  
#Calculate median overall and potential scores across all players
median_overall <- df_clean %>%
  summarise(median_value = median(Overall.Score, na.rm = TRUE)) %>%
  pull(median_value)  # Converts it to a vector

median_potential <- df_clean %>%
  summarise(median_value = median(Potential.Score, na.rm = TRUE)) %>%
  pull(median_value)


#Assign performance categories, calculating the share of each category per League
df_clean<-df_clean %>% mutate( 'Performance Category' = case_when(
      Overall.Score >= median_overall & Potential.Score >=median_potential ~ "Emerging High Performers",  # "High Overall, High Potential"
      Overall.Score >= median_overall & Potential.Score < median_potential ~ "High Peak Performers", # "High Overall, Low Potential"
      Overall.Score < median_overall & Potential.Score >= median_potential ~ "Future High Performers", #  "Low Overall, High Potential"
      Overall.Score < median_overall & Potential.Score < median_potential ~ "Limited Performance" # "Low Overall, Low Potential"
    )) %>%
  group_by(LeagueRanked) %>%
  mutate(Total_Count = n()) %>%
  
  group_by(LeagueRanked, `Performance Category`) %>%
  mutate(Count = n()) %>% 
  
   mutate(Percentage = round((Count / Total_Count) * 100, 0)) %>%
  group_by(LeagueRanked) %>%
  mutate(
    x_pos = case_when(
      `Performance Category` %in% c("Emerging High Performers", "High Peak Performers") ~ 90,
      `Performance Category` %in% c("Future High Performers", "Limited Performance") ~ 50
    ),
    y_pos = case_when(
      `Performance Category` %in% c("Emerging High Performers", "Future High Performers") ~ 90,
      `Performance Category` %in% c("High Peak Performers", "Limited Performance") ~ 60
    )
  ) %>%
  
  ungroup()

avg_wage_per_league <- df_clean %>%
  group_by(LeagueRanked,League_Rank) %>%
  summarise(Average_Wage = mean(Wage, na.rm = TRUE) / 1000) 

head(df_clean)
```

## EDA

### Graph 1. Amount of players across Top Football Leagues

Sorted by League Rank

```{r,  warning=FALSE,message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
#install.packages("ggsci")
library(ggsci)
 ggplot(df_clean) +
  aes(x = reorder(LeagueRanked, -League_Rank), fill = LeagueRanked) +  # Sort leagues by rank
  geom_bar( color = "666666") +
  coord_flip() +
  scale_fill_observable()+
  labs(title = "Amount of Players", x = "Count", y = "League") +
   theme( legend.position = "none",plot.title = element_text(size = 16))
  

```

When looking at the player distribution across the top football leagues, it's clear that most leagues have a similar number of players, usually between **400 and 500**. However, there are noticeable differences, especially in the **Argentine** and **Brazilian** leagues.

The **Argentine Primera División** has **28 clubs**, which is more than the usual **20 clubs** found in most top leagues. This larger number of teams naturally leads to a higher total number of registered players.

In contrast, the **Campeonato Brasileiro Série A** has **20 clubs**, similar to many major European leagues.Even with the same number of clubs, differences in **player registration rules** and **squad management** can cause variations in the total number of players across leagues.

### Graph 2. Overall Score and Age Distribution Across Top Football Leagues

```{r,  fig.width=14, fig.height=7, warning=FALSE, message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
#install.packages("patchwork")

library(patchwork)
# 📊 **Graph 1
p1 <- ggplot(df_clean) +
  aes(x = Overall.Score, y = reorder(LeagueRanked, -League_Rank), fill = LeagueRanked) +  # Sort leagues by rank
  geom_boxplot( color = "666666") +
 scale_fill_observable()+
  labs(title = "Score Distribution", x = "Overall Score", y = "League") +
   theme( legend.position = "none",plot.title = element_text(size = 14))
  
  

# 📊 **Graph 2
p2 <- ggplot(df_clean) +
  aes(x = Age, y=reorder(LeagueRanked, -League_Rank), fill = LeagueRanked) +  # Sort leagues by rank
  geom_boxplot( color = "666666") +
  scale_fill_observable()+
  labs(title = "Age Distribution", x = "Age", y = NULL) +
  theme(axis.text.y = element_blank(), legend.position = "none",plot.title = element_text(size = 14))


# 🖼 **Combine the graphs horizontally**
p1 + p2 + plot_layout(ncol = 2)+ 
  plot_annotation(title = "Overall Score and Age Distribution")

```

The **score distribution** reflects the **league rankings**, with top leagues like the **Premier League**, **La Liga**, and **Bundesliga** having players with **higher average scores**, showing that the best talent is often found in these leagues. On the other hand, lower-ranked leagues have a **wider range of scores**, meaning there's more variety in player performance.

The **age distribution** also shows that **top leagues are focused on signing young, talented players**, as they have **younger average ages** compared to other leagues. This suggests that top leagues are always on the lookout for the next generation of football stars to stay competitive.

### Graph 3. Player Performance Distribution Across Top Football Leagues

This graph shows the distribution of football players based on their **Overall Score** (*x-axis*) and **Potential Score** (*y-axis*) across the top 10 football leagues.

Players are categorized into four **performance groups**, defined using the **median Overall Score** and **median Potential Score**, calculated across all players:

-   **🟢 Emerging High Performers (light green):**\
    Players with **high current performance and high potential**, positioned **above both medians**.

-   **🟢 Future High Performers (dark green):**\
    Players with **lower current performance but high potential**, falling **below the median for Overall Score** but **above the median for Potential Score**.

-   **⚫ High Peak Performers (gray):**\
    Players with **high current performance but limited growth potential**, identified as having an **Overall Score above the median** but **Potential Score below the median**.

-   **🔴 Limited Performance (red):**\
    Players with **low current performance and low potential**, positioned **below both medians**.

```{r, warning=FALSE, fig.width=14, fig.height=7, message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"

ggplot(df_clean %>% filter(!is.na(`Performance Category`))) + aes(x = Overall.Score, y = Potential.Score, color = `Performance Category`)+ geom_jitter(alpha = 0.6) + scale_color_manual(values = c( "Emerging High Performers" = "#91D1C2B2","High Peak Performers" = "gray20","Future High Performers" ="#1b9e77","Limited Performance" = "666666")) +# Horizontal median line for each league
   geom_vline(aes(xintercept = median_overall), color = "gray40", linetype = "dashed", linewidth = 1) +
   
   # Vertical median line for each league
   geom_hline(aes(yintercept = median_potential), color = "gray40", linetype = "dashed", linewidth = 1) + geom_text(data = df_clean%>% group_by(LeagueRanked,`Performance Category`)%>% distinct(LeagueRanked, `Performance Category`, .keep_all = TRUE) , aes(x = x_pos, y = y_pos, label = paste0(Percentage, "%")), inherit.aes = FALSE, size = 4.5,hjust = 0.5) + facet_wrap(vars(reorder(LeagueRanked, League_Rank)), ncol = 4) + labs(title = "Where Future Talants Emerge", x = "Overall Score (Mean = 70)", y = "Potential Score (Mean = 75)") + theme(legend.position = "right",legend.key.height = unit(1, "cm"), plot.title = element_text(size = 14))+ guides(color = guide_legend(override.aes = list(size = 12)))

```

Top European leagues like the **Premier League (England)**, **La Liga (Spain)**, and **Bundesliga (Germany)** have a **higher percentage of Emerging High Performers** attract players who are already performing at a high level while still having room for growth.

In contrast, leagues such as the **Süper Lig (Turkey)** and **Liga Profesional de Fútbol (Argentina)** show a **higher concentration of Limited Performance players** (**63%** and **51%**, respectively).

Leagues like the **Eredivisie (Netherlands)** and **Primeira Liga (Portugal)** stand out for their **strong representation of Future High Performers** (**29%** and **23%**). These leagues are well-known for **developing young talents** who often move on to larger clubs in top-tier competitions.

Meanwhile, leagues such as **Serie A (Italy)** and **Ligue 1 (France)** display a **more balanced distribution** across all performance categories. This reflects a healthy mix of **Veteran stars**, **Future talents**, and **Squad players**.

### Graph 4. Contract Duration vs Overall Score Across Top Football Leagues

This graph illustrates the relationship between players' Overall Score (x-axis) and their Contract Duration in years (y-axis) across the top 10 football leagues. Each black dot represents an individual player, while the red line indicates the trend line (linear regression), showing the general direction of the relationship.

```{r,  fig.width=14, fig.height=7, warning=FALSE,  message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"

ggplot(df_clean) +
  aes(x = Overall.Score, y = Contract.Duration, fill = LeagueRanked) +  
  geom_jitter(alpha = 0.6, shape = 21, size = 2, color="gray60") +    
  scale_fill_observable() +                                           
  geom_smooth(method = "lm", color = "gray10", linetype = "solid", se = FALSE) + 
  facet_wrap(vars(reorder(LeagueRanked, League_Rank)), ncol = 5L) +
  labs(
    title = "Contract Duration vs Overall Score Across Top Football Leagues",
    x = "Overall Score",
    y = "Contract Duration"
  ) +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 14)
  )
```

While there's a general trend of longer contracts for higher-rated players, the strength of this relationship varies by league. Top leagues seem to reward talent with longer commitments, while others may have less flexibility due to financial or structural factors.

### Graph 5. Wage Distribution Across Top Football Leagues

Have you ever been curious about how much football players earn?

This graph shows the average weekly salary (in thousands of euros) for players in the top 10 football leagues.

```{r,  fig.width=14, fig.height=7, warning=FALSE,  message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
# Plotting average wage as bars
ggplot(avg_wage_per_league) +
  aes(x = Average_Wage, y = reorder(LeagueRanked, -League_Rank), fill = LeagueRanked) +
  geom_col(color = "666666", width = 0.7) +  # Bar plot with custom color
  scale_fill_observable() +
  labs(
    title = "Average Wage Across Top Football Leagues",
    x = "Average Wage  (€K)", 
    y = "League"
  ) +
  theme(
    legend.position = "none",
   
  )+
   geom_text(aes(label = paste0(round(Average_Wage, 1), "K")), 
            hjust = -0.1, size = 4, color = "black")

```

The **Premier League (England)** leads with **€71.6K**, nearly **double** that of the next two highest-paying leagues---**Serie A (Italy)** at **€40.5K** and **La Liga (Spain)** at **€37.8K**.

Compared to leagues like the **Eredivisie (Netherlands)** (**€8K**) and **Primeira Liga (Portugal)** (**€8.7K**), Premier League players earn almost **9 times more**.

Interestingly, the **Seria A(Brazil)** and **Süper Lig (Turkey)** show relatively **high average wages** compared to leagues with similar competitive levels.

### Graph 6. Wage Distribution by Role

```{r,  fig.width=14, fig.height=7, warning=FALSE, message=FALSE }
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"

avg_wage_per_role <- df_clean %>%
  group_by(Role) %>%
  summarise(Average_Wage = mean(Wage, na.rm = TRUE) / 1000) 


ggplot(avg_wage_per_role) +
  aes(x = Average_Wage, y =  reorder(Role,Average_Wage), fill = Role) +
  geom_col(color = "666666", width = 0.7) +  # Bar plot with custom color
  scale_fill_manual(values = c(
    "Winger" = "#FF6FB5",          
    "Side Midfielder" = "#8BD3E6", 
    "Other" = "gray80"             
  ))+
  labs(
    title = "Average Wages by Football Roles",
    x = "Average Wage  (€K)", 
    y = "Role"
  ) +
  theme(
    legend.position = "none",
  )

```

Wingers (pink) earn the most, likely due to their game-changing impact in attack.

Side Midfielders (light blue) earn the least, as their role is often less decisive in scoring or defense.

Strikers, Midfielders, and Defenders fall in between, reflecting their varying influence on the game.

### Graph 7. Salary Trends and Tactical Priorities Across Leagues

Let's combine the last two charts to see if Wingers consistently earn the most and Side Midfielders the least across all leagues.

```{r,  fig.width=14, fig.height=7, warning=FALSE,  message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
avg_wage_by_league_role <- df_clean %>%
  group_by(LeagueRanked, League, Role, League_Rank) %>%
  summarise(Average_Wage = round(mean(Wage, na.rm = TRUE)/1000,0)) %>%
  group_by(LeagueRanked) %>%
  mutate(Role_Rank = dense_rank(desc(Average_Wage))) %>%
  ungroup()%>%
  mutate(Highlight = case_when(
    Role == "Winger" ~ "Winger",
    Role == "Side Midfielder" ~ "Side Midfielder",
    TRUE ~ "Other"  
  ))

# Plotting the average wage
library(tidytext)  # For reorder_within()


# Plot with Roles Ordered Separately for Each League
ggplot(avg_wage_by_league_role%>%filter(League %in% c("Bundesliga (Germany)", "Serie A (Italy)", "La Liga (Spain)", 
                        "Ligue 1 (France)", "Premier League (England)"))) +
  aes(x = reorder_within(Role, -Average_Wage, LeagueRanked),  
      y = Average_Wage, 
      fill = Highlight) +  
  geom_col(width = 0.7) +  
   scale_fill_manual(values = c(
    "Winger" = "#FF6FB5",          
    "Side Midfielder" = "#8BD3E6", 
    "Other" = "gray80"             
  ))+
  facet_wrap(vars(reorder(LeagueRanked, League_Rank)), ncol = 5, scales = "free_x") +  # Allow independent ordering
  scale_x_reordered() +  # Fix axis labels after reordering
  labs(
    title = "Average Wage by Role in Top Football Leagues",
    x = "Role",
    y = "Average Wage (€K)"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "none"
  )
```

Salary trends reflect not only player performance but also the **tactical priorities** of each league. **Wingers** are highly valued in leagues that emphasize fast, attacking football, such as the **Premier League**, while leagues like **La Liga** tend to reward **goal-scoring central roles**.

Despite these differences, **side midfielders consistently receive the lowest wages** across all leagues, highlighting a trend where their contributions are less financially recognized.

Interestingly, the **Bundesliga** stands out with a more **balanced wage distribution**, suggesting a strategic focus on **team-oriented play** rather than heavy reliance on specific star positions.

```{r,  fig.width=14, fig.height=7, warning=FALSE,message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"

ggplot(avg_wage_by_league_role%>%filter(!League %in% c("Bundesliga (Germany)", "Serie A (Italy)", "La Liga (Spain)", 
                        "Ligue 1 (France)", "Premier League (England)"))) +
  aes(x = reorder_within(Role, -Average_Wage, LeagueRanked),  
      y = Average_Wage, 
      fill = Highlight) +  
  geom_col(width = 0.7) +  
   scale_fill_manual(values = c(
    "Winger" = "#FF6FB5",          
    "Side Midfielder" = "#8BD3E6", 
    "Other" = "gray80"             
  ))+
  facet_wrap(vars(reorder(LeagueRanked, League_Rank)), ncol = 5, scales = "free_x") +  # Allow independent ordering
  scale_x_reordered() +  # Fix axis labels after reordering
  labs(
    title = "Average Wage by Role in Top Football Leagues",
    x = "Role",
    y = "Average Wage (€K)"
  ) +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), 
    strip.text = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    legend.position = "none"
  )
```

Compared to the top 5 European leagues, only the **Süper Lig (Turkey)** shows a similar pattern, with **wingers** earning the most and **side midfielders** the least. **Argentina's Liga Profesional** has a **balanced wage distribution**, much like the **Bundesliga**, reflecting a focus on **team-wide value** rather than specific roles.

In **Brazil's Série A**, **defensive roles** earn wages comparable to attackers, highlighting the **importance of defensive contributions**.

Meanwhile, both **Portugal's Primeira Liga** and the **Netherlands' Eredivisie** show **lower, more balanced wages**, which aligns with their emphasis on **youth development** and maintaining **financial sustainability** within clubs.

## Multi-Linear Regression Models

Lets run Multi-Linear Regression Models for the top 5 leagues to understand how different player performance metrics and characteristics influence their wages.

### Prepere Data for OLS

To prepare the data for OLS regression, I created dummy variables for player roles, converted wages to thousands of Euros, and removed unnecessary columns and missing data. I also removed highly correlated variables to avoid multicollinearity, keeping the dataset clean for analysis.

```{r, warning=FALSE, message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
df_ols <- df_clean %>%
  mutate(
    Winger = ifelse(Role == "Winger", 1, 0),
    Striker = ifelse(Role == "Striker", 1, 0),
    Central_Midfielder = ifelse(Role == "Central Midfielder", 1, 0),
    Central_Defender = ifelse(Role == "Central Defender", 1, 0),
    Side_Defender = ifelse(Role == "Side Defender", 1, 0),
    Wage = Wage / 1000
  ) %>%
    select(
    -c(Player, LeagueRanked, `Performance Category`, Role, Total_Count,Count,Percentage,x_pos,y_pos,Growth,Base.Stats,Weight_kg,League_Rank,Height_cm                  ,Jumping,Heading.Accuracy ,Sliding.tackle,Left_Foot,Contract.Duration,Penalties )) %>%
  drop_na()


colnames(df_ols)

```

In the next code snippet, I run an Ordinary Least Squares (OLS) regression. Before that, I apply the Recursive Feature Elimination (RFE) method to select the top 18 most important variables that influence player wages. This method helps reduce dimensionality and improve model performance by iteratively removing less significant features. After fitting the initial regression model, I keep only the variables that are statistically significant (p-value \< 0.05). This process resembles backward elimination because it systematically removes insignificant variables, although the primary selection is done through RFE.

```{r, warning=FALSE, message=FALSE}
#| echo: true
#| code-fold: true
#| code-summary: "expand for full code"
#install.packages("caret")
#install.packages("MASS")
library(MASS) 
library(caret)
library(dplyr)

run_RFE <- function(league_name, corr_threshold = 0.75, p_value_threshold = 0.05) {

  data_league <- df_ols %>% filter(League == league_name) %>% dplyr::select(-"League")
  X_filtered <- data_league %>% dplyr::select(-c("Wage"))
  y <- data_league$Wage

  corr_matrix <- cor(X_filtered, use = "pairwise.complete.obs")
  high_corr_vars <- findCorrelation(corr_matrix, cutoff = corr_threshold, names = TRUE, exact = TRUE)
  X_filtered <- X_filtered %>% dplyr::select(-all_of(high_corr_vars))
  
  control <- rfeControl(functions = lmFuncs, method = "cv", number = 10)
  rfe_model <- rfe(X_filtered, y, sizes = 18, rfeControl = control)
  best_vars <- predictors(rfe_model)
  
  data_league <- data_league %>% dplyr::select(Wage, Value, all_of(best_vars))
  final_formula <-as.formula(paste("Wage~", paste(best_vars,collapse=" + ")))
  final_model <- lm(final_formula, data = data_league)
  model_summary<-summary(final_model)
    
  significant_vars <- names(which(model_summary$coefficients[, 4] < p_value_threshold))
  significant_vars <- significant_vars[significant_vars != "(Intercept)"]  
  
  if (length(significant_vars) > 0) {
    final_model <- lm(as.formula(paste("Wage ~", paste(significant_vars, collapse = " + "))), 
                      data = data_league)
    model_summary <- summary(final_model)
  } 

  return(list(data_league=data_league, model_summary=model_summary))
}

# Apply to each league
results_England <- run_RFE("Premier League (England)")
results_Germany <- run_RFE("Bundesliga (Germany)")
results_Italy   <- run_RFE("Serie A (Italy)")
results_Spain   <- run_RFE("La Liga (Spain)")
results_France  <- run_RFE("Ligue 1 (France)")

```

### MLR for Premier League (England)

```{r}

print(results_England$model_summary)

```

### MLR for La Liga (Spain)

```{r}
print(results_Spain$model_summary)
```

### MLR for Bundesliga (Germany)

```{r}
print(results_Germany$model_summary)
```

### MLR for Serie A (Italy)

```{r}
print(results_Italy$model_summary)
```

### MLR for Ligue 1 (France)

```{r}
print(results_France$model_summary)
```

The OLS regression results across the top 5 European leagues reveal that International Reputation is the strongest and most consistent factor influencing player wages. However, this likely reflects a two-way relationship, where playing in top leagues boosts a player's international reputation, which in turn is linked to higher wages.

Potential Score and Reactions also play significant roles, indicating the importance of both future growth potential and on-field responsiveness.
